A Neural, Parallel Theorem Prover*

### 2019-05-10 papers/FroCoS19/paper.tex
added ML section

\documentclass[runningheads]{llncs}
\usepackage{graphviz}
\usepackage{hyperref}
\newcommand{\lerna}{\textsc{Lerna}}
\newcommand{\z}[1]{\textsc{Z3}}
\newcommand{\mizarlarge}{\textit{M40k}}
\newcommand{\mizarsmall}{\textit{M2k}}
\begin{document}
\title{A Neural, Parallel Theorem Prover*}
\author{
	Michael Rawson*\orcidID{0000-0001-7834-1567}\inst{1} \and
	Giles Reger\inst{2}
}
\institute{
	University of Manchester, UK\\
	\email{michael@rawsons.uk}\and
	University of Manchester, UK
}
\maketitle

@@

\subsection{Data Collection}
A large dataset of satisfiable and unsatisfiable goals were collected by running the unguided prover on the \mizarlarge{} dataset for 10 seconds.
As soon as the prover determines the satisfiability of any sub-goal, the formula it represents and its status is recorded.
This resulted in 18,340 unsatisfiable examples and 1,845,267 satisfiable examples, occupying 6GB of disk space.
The dataset is very imbalanced (due to a combination of weakening rules producing a large number of trivially-satisfiable examples, and to prover termination after after the goal is shown to be unsatisfiable), at a ratio of around 100:1.

\subsection{Translation to Graphs}
Wang et al.~\cite{formula-graph} give a translation from higher-order formulae to directed graphs, and a similar scheme is used here.
Constants, function symbols, predicate symbols, and bound variables are given their own node.
Application of functions and predicates to arguments are represented as an ``application node'' with two children: the symbol node and an ``argument list'' node representing the list of arguments.
Propositional connectives and equality have the obvious representation, while quantifiers have two children: the variable they bind and their sub-formula.
Common sub-graphs are merged.

An example formula's translation from
\[
	\forall X, \left[p\left(f(c), X\right) \wedge \lnot\left(f(c) = d\right)\right]
\]
is shown in figure...

\begin{figure}
\end{figure}

\subsection{Augmentation}
\subsection{Neural Architecture}

\subsection{Neural Network Throughput}
To measure network throughput, the network evaluated \(\approx\) 500MB of pre-recorded data with various batch sizes --- batch size represents a tradeoff between latency of evaluations and parallelism opportunity for a GPU.
Mean throughput (measured in evaluations/second) for each batch size is shown in Table \ref{table:network-throughput}, representing the total number of tableaux sub-goals evaluated in a single second.
As expected, neural network throughput for this task increases rapidly when allowed to process inputs in batches.
In practice a larger batch size was more helpful for proof search, suggesting that a small amount of latency does not significantly affect search in the long term.

\begin{table}
	\caption{Mean neural network throughput versus batch size: evaluations per second.}
	\centering
	\begin{tabular}{c | c}
		\textbf{Batch Size} & \textbf{Mean Throughput (evals./sec)}\\
		\hline
		1 & 254.2\\
		8 & 887.9\\
		32 & 2,128.3\\
		128 & 3,015.0\\
	\end{tabular}
	\label{table:network-throughput}
\end{table}

>>>>

\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{subcaption}
\newcommand{\lerna}{\textsc{Lerna}}
\newcommand{\z}[1]{\textsc{Z3}}
\newcommand{\mizarlarge}{\textit{M40k}}
\newcommand{\mizarsmall}{\textit{M2k}}
\begin{document}
\title{A Neural, Parallel Theorem Prover*}
\author{
	Michael Rawson*\orcidID{0000-0001-7834-1567}\inst{1} \and
	Giles Reger\inst{2}
}
\institute{
	University of Manchester, UK\\
	\email{michael@rawsons.uk}\and
	University of Manchester, UK
}
\maketitle

@@

\subsection{Data Collection}
A large dataset of satisfiable and unsatisfiable goals were collected by running the unguided prover on the \mizarlarge{} dataset for 10 seconds.
As soon as the prover determines the satisfiability of any sub-goal, the formula it represents and its status is recorded.
This resulted in 18,340 unsatisfiable examples and 1,845,267 satisfiable examples, occupying 6GB of disk space.
The dataset is very imbalanced (due to a combination of weakening rules producing a large number of trivially-satisfiable examples, and to prover termination after the goal is shown to be unsatisfiable), at a ratio of around 100:1.

\subsection{Translation to Graphs}
Wang et al.~\cite{formula-graph} give a translation from higher-order formulae to directed graphs, and a similar scheme is used here.
Constants, function symbols, predicate symbols, and bound variables are given their own node.
Application of functions and predicates to arguments are represented as an ``application node'' with two children: the symbol node and an ``argument list'' node representing the list of arguments.
Propositional connectives and equality have the obvious representation, while quantifiers have two children: the variable they bind and their sub-formula.

To produce an input graph from a formula \(F\), the formula is first parsed into an abstract syntax tree.
Common sub-trees up to \(\alpha\)-equivalence are merged, then the resulting directed acyclic graph has any named-symbol nodes replaced with an opaque, nameless label such as "predicate" or "variable" --- since distinct symbols remain as distinct nodes under this scheme, no information is lost other than the natural-language semantics of the symbol name.
In practice, undirected graphs improved model performance so the graph is made undirected before encoding node labels as one-hot inputs to produce the final input graphs.

An example formula's translation from
\[
	F = \forall X, \left[p\left(f(c), X\right) \wedge \lnot\left(f(c) = d\right)\right]
\]
is shown in Figure \ref{figure:translation}.

\begin{figure}
	\centering
	\begin{subfigure}{.4\textwidth}
		\centering
		\includegraphics[width=\linewidth]{translation-ast}
		\caption{The AST for \(F\).}
	\end{subfigure}
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\linewidth]{translation-graph}
		\caption{De-duplicated DAG.}
	\end{subfigure}
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\linewidth]{translation-nameless}
		\caption{Nameless DAG.}
	\end{subfigure}
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\linewidth]{translation-undirected}
		\caption{Final result.}
	\end{subfigure}
	\caption{The translation process from \(F\) to a graph as seen by the neural network.}
	\label{figure:translation}
\end{figure}

\subsection{Augmentation}
One possible solution~\cite{imbalanced} to the problem of classification on imbalanced domains is to synthesize new data for under-represented classes --- in this case unsatisfiable formulae --- from existing data by augmenting it.
A classic example is augmenting image data by cropping, flipping or adding noise to existing images.

There are many possible ways to augment formulae graphs.
For this prototype, a simple approach is taken in which a small number of nonsense formulae are added to the graph by randomly adding nodes/edges where appropriate.
This approach has the advantage of exposing the network to ``noise'' such as additional axioms which might well occur in practice, but if the network is adequately capable of filtering these then no new formulae are actually seen.

\subsection{Neural Architecture}
TODO set fixed architecture and report on this.

\subsection{Implementation}
TODO discuss implementation with PyTorch~\cite{pytorch} and its graph-processing extension PyTorch Geometric~\cite{pytorch-geometric}.
